### 2011-06-08 draft2.tex
PDF Update

\section{Example}
Consider the following program (the integer after a data constructor is its arity):
\begin{verbatim}
data List = Nil 0
          | Cons 2;;

@@

copy ::: a:{x : True} -> {y:True};;
\end{verbatim}
The contract states that if we give to copy a crash-free argument $x$, copy will return a crash-free result $y$.
Note that [Just BAD] is not crash-free in our sense, so we can't say anything about it. TODO (or can we?)

\begin{verbatim}
data Nat = Zero 0
         | Succ 1;;

@@

add ::: a:{x : True} -> {y:True} -> {z: True};;             -- C1
add ::: a:{x : isZero x} -> {y:isZero y} -> {z: True};;     -- C2
add ::: a:{x : isZero x} -> {y:isZero y} -> {z: isZero z};; -- C3
\end{verbatim}

As of today, we can prove C1 and C2 but not C3.
Note that we can prove C2 only since the modification of $S_4$ in $\dtrans{}$.

@@

>>>>

\section{Example}

\subsection{Semantics}
Consider the following program (the integer after a data constructor is its arity):
\begin{verbatim}
data List = Nil 0
          | Cons 2;;

@@

copy ::: a:{x : True} -> {y:True};;
\end{verbatim}
The contract states that if we give to copy a crash-free argument $x$, copy will return a crash-free result $y$.
Note that [Just BAD] is not crash-free in our sense, so we can't say anything about it. TODO

\subsection{Arguments order}
\begin{verbatim}
data Nat = Zero 0
         | Succ 1;;

@@

add ::: a:{x : isZero x} -> {y:isZero y} -> {z: isZero z};;
\end{verbatim}

@@

A lemma stating (+) commutativity would be helpful here.

\subsection{Multiple contracts - Take 1}
We suppose defined datatypes for Nat and Bool, just as in the example above.

\begin{verbatim}
add a b = case a of
  | Zero -> b
  | Succ x -> Succ (add x b);;

mult a b = case a of
  | Zero -> Zero
  | Succ x -> add b (mult x b);;

notZero x = case x of
  | Zero -> False
  | Succ a -> True;;

notZero ::: a:{x:True} -> {y: True};;
add  ::: a:{x: True} -> b:{y: True} -> {z: True};;
add  ::: a:{x: notZero x} -> b:{y: True} -> {z: True};;
mult ::: a:{x: notZero x} -> b:{y: notZero y} -> {z: True};;
\end{verbatim}

Equinox diverges when we try to check mult's contract.

The reason stems from that $mult~(Succ~x)~b$ calls recursively
$mult~x~b$ The induction hypothesis says that if $a$ is nonzero and
$b$ is nonzero then the result is $\ok$. But in this case $x$ may be
zero, so the induction hypothesis does not apply immediately. Now, the
theorem prover could do a split on $x$ (either Zero or Succ$\_$) and
in one case the result should follow from the induction hypothesis,
but in the other case we are stuck! Because we don't actually know
that $multp~0~b$ is $\ok$ ($multp$ is the recursive version of the
function, whose contract we assumed but no further defining
equations!)

A possible fix, which is only possible when the function is total is
to write $$f ::: a_1:\{x \mid p_1(x) \} \to \cdots \to a_n:\{x \mid
p_n(x) \}$$ as $$f ::: \ok \to \cdots \to \ok \to \{x \mid
if~(\bidWedge_{i < n} p_i(a_i))~then~p_n(x)~else~True)$$ And know we
can prove $mult$'s contract, but it's not very satisfying.
